---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
input <- read.csv("activity.csv")
input$date <- as.Date(input$date, format="%Y-%m-%d")
head(input)
```




## What is mean total number of steps taken per day?
```{r}
aggr <- aggregate(input[c("steps")], by =list(input$date), FUN=sum, na.rm=TRUE)
aggr
hist(aggr$steps, main = "The Total number of steps taken each day", 
     xlab = "steps", breaks = 50)

```
```{r}
aggrmean <- aggregate(input[c("steps")], by=list(input$date), FUN=mean,
                  na.rm=TRUE)
aggrmedian <- aggregate(input[c("steps")], by=list(input$date), FUN=median,
                  na.rm=TRUE)
meanMedian <- data.frame(date=aggrmean$Group.1, Mean=aggrmean$steps,    Median=aggrmedian$steps)
names(meanMedian) <- c("date", "mean", "median")
meanMedian
```

Note: The mean and median columns above may produce NA's as these days have all the steps as NA.


## What is the average daily activity pattern?

```{r}
aggrInterval <- aggregate(input[c("steps")], by = list(input$interval), FUN=mean, na.rm=TRUE)
names(aggrInterval) <- c("interval", "steps")
plot(aggrInterval$interval, aggrInterval$steps, type= "l", main = "Average number of steps taken", xlab = "Intervals" , ylab="Average steps" )
head(aggrInterval, n=50)

```

The Interval that contains the maximum number of average steps is
```{r}
aggrInterval$interval[which(aggrInterval$steps== max(aggrInterval$steps))]

```


## Imputing missing values

The total number of missingvalues include
```{r}
sum(is.na(input$steps))
```



using the mean of the 5-minutes interval for imputing missing values
```{r}
newdata <- input
for(i in 1:nrow(input)) {
    if(is.na(newdata$steps[i])) {
        newdata$steps[i] <- aggrInterval$steps[which(aggrInterval$interval == newdata$interval[i])]
    }    
}
```

Calculating the sum of steps taken for the newly created dataset.
```{r}
newsum <- aggregate(newdata[c("steps")], by =list(newdata$date), FUN=sum, na.rm=TRUE)
newsum
hist(newsum$steps, main = "The Total number of steps taken each day", 
     xlab = "steps", breaks = 50)

```
* Imputing missing values has the effect of normalizing the distribution as it can be seen from the histigram

Calculating mean and median for the new Dataset.
```{r}
newmean <- aggregate(newdata[c("steps")], by=list(newdata$date), FUN=mean,
                  na.rm=TRUE)
newmedian <- aggregate(newdata[c("steps")], by=list(newdata$date), FUN=median,
                  na.rm=TRUE)
newmeanMedian <- data.frame(date=newmean$Group.1, Mean=newmean$steps,    Median=newmedian$steps)
names(newmeanMedian) <- c("date", "mean", "median")
newmeanMedian

```

* The mean and median values of the newly created dataset doesn't deviate from the first part of the assignment.


## Are there differences in activity patterns between weekdays and weekends?

Calculating whether wekday or weekend. and plotting the interval mean plot.
```{r}
week <- vector(mode="character", length = nrow(input))
for(i in 1:nrow(input)) {
    date <- as.Date(input$date[i],format="%Y-%m-%d" )
    if(weekdays(date)=="Saturday" || weekdays(date)=="Sunday") {
        week[i] <- "weekend"
    
    }
    else
        week[i] <- "weekday"
    
}
updatedframe <- data.frame(input, week)
par(mfrow = c(2,1))
day <- subset(updatedframe, updatedframe$week=="weekday")
end <- subset(updatedframe, updatedframe$week=="weekend")
dayaggr <- aggregate(day[c("steps")], by=list(day$interval), FUN=sum,na.rm=TRUE)
endaggr <- aggregate(end[c("steps")], by=list(end$interval), FUN=sum,na.rm=TRUE)
plot(dayaggr$Group.1,dayaggr$steps , type="l", ylab = "Number of steps", main = "Weekday", xlab="Interval")
plot(endaggr$Group.1,endaggr$steps , type="l", ylab = "Number of steps", main = "Weekend", xlab = "Interval")



```


